\section{Method}\label{sec:method}

The hockey environment exposes a low-dimensional but fully continuous action space, 
which makes discrete-action methods such as DQN~\cite{mnih2015:DQN}
inapplicable and on-policy methods such as PPO~\cite{SchulmanEtAl2017:PPO} less sample-efficient
due to their inability to reuse past experience.
We therefore focus on two off-policy, continuous-control baselines that represent complementary
design philosophies: TD3~\cite{fujimoto2018:TD3}, which optimizes a deterministic policy with
explicit bias-correction mechanisms, and SAC~\cite{haarnoja2018soft}, which optimizes a
stochastic policy through entropy maximization.
Both are well-established benchmarks in continuous control, offer strong sample efficiency via
experience replay, and their contrasting approaches to exploration and value estimation make a
direct empirical comparison meaningful on the same task.

Both algorithms share a common backbone: a replay buffer $\mathcal{D}$ stores transitions
$(s_t, a_t, r_t, s_{t+1}, d_t)$, and gradient updates are performed on minibatches sampled from
$\mathcal{D}$.
An actor $\pi$ selects actions while one or more critics $Q_\theta$ estimate action-values; both
components maintain slowly updated target networks updated via Polyak averaging,
\begin{equation}
  \theta' \leftarrow \tau\,\theta + (1-\tau)\,\theta', \qquad
  \phi'   \leftarrow \tau\,\phi  + (1-\tau)\,\phi'.
  \label{eq:polyak}
\end{equation}
TD3 (Section~\ref{sec:td3}) uses a \emph{deterministic} actor with three targeted stabilization
mechanisms; SAC (Section~\ref{sec:sac}) uses a \emph{stochastic} actor augmented with entropy
regularization.
Both are further extended with the improvements described in Section~\ref{sec:improvements}.


\newpage


\subsection{Twin Delayed Deep Deterministic Policy Gradient (TD3)}
\label{sec:td3}

TD3~\cite{fujimoto2018:TD3} is a model-free, off-policy actor--critic algorithm that extends
DDPG~\cite{lillicrap2015:DDPG} by explicitly addressing overestimation bias and policy update
instability in continuous action domains.
The actor encodes a \emph{deterministic} policy $a = \pi_\phi(s)$ and is trained to maximize the
expected discounted return
\begin{equation}
  J(\phi) = \mathbb{E}\!\left[\sum_{t=0}^{\infty} \gamma^t\, r(s_t, \pi_\phi(s_t))\right].
\end{equation}
Following the deterministic policy gradient theorem~\cite{silver2014:DPG}, the gradient of this
objective with respect to the actor parameters is
\begin{equation}
  \nabla_\phi J(\phi)
    \approx \mathbb{E}_{s \sim \mathcal{D}}\!\left[
      \nabla_a Q_{\theta}(s,a)\big|_{a=\pi_\phi(s)}\,
      \nabla_\phi \pi_\phi(s)
    \right].
\end{equation}
This allows the actor to be updated by simply back-propagating the critic gradient through the
deterministic action.
While this yields highly informative gradient signal in continuous domains, DDPG-style methods
are acutely sensitive to critic inaccuracies: whenever the critic overestimates the value of
certain actions, the actor exploits these errors and converges to suboptimal or unstable
behaviour~\cite{fujimoto2018:TD3}.
TD3 addresses this with three interacting mechanisms.

\textbf{Clipped double Q-learning.}
To counteract overestimation bias, TD3 maintains two independent critics $Q_{\theta_1}$ and
$Q_{\theta_2}$ that are trained in parallel but receive no shared parameters.
When computing Bellman targets, the minimum of the two target critics is used,
\begin{equation}
  y = r + \gamma(1-d)\min_{i=1,2} Q_{\theta_i'}(s', \tilde{a}),
  \label{eq:td3target}
\end{equation}
where $d$ indicates episode termination and $\theta_i'$ are the target network parameters updated
via~\eqref{eq:polyak}.
Taking the minimum introduces a conservative bias that counteracts the systematic overestimation
inherent in maximizing over noisy function approximators.
Both critics are then independently updated by minimizing the mean squared Bellman error
$(y - Q_{\theta_i}(s, a))^2$ over minibatches from $\mathcal{D}$.
Because the target $y$ is shared but gradients flow independently, the two critics diverge
slightly, which is precisely what prevents correlated overestimation.

\textbf{Target policy smoothing.}
A deterministic policy is particularly vulnerable to narrow, erroneous peaks in the learned value
function: if the critic assigns a spuriously high value to a small region of action space, the
actor will immediately collapse onto that region.
TD3 regularizes against this by perturbing the target action with clipped Gaussian noise before
evaluating the Bellman target,
\begin{equation}
  \tilde{a} = \operatorname{clip}\!\left(\pi_{\phi'}(s') + \epsilon,\;a_{\min},\;a_{\max}\right),
  \quad \epsilon \sim \operatorname{clip}\!\left(\mathcal{N}(0,\sigma),\;-c,\;c\right).
\end{equation}
By forcing the backup to average over a small neighbourhood of the target action, the critics are
encouraged to learn a smoother value landscape, making individual erroneous peaks less
exploitable by the actor~\cite{fujimoto2018:TD3}.
The noise standard deviation $\sigma$ and clip range $c$ are hyperparameters that control the
effective radius of this smoothing.

\textbf{Delayed policy updates.}
Even with conservative targets and smoothed backups, updating the actor at every critic step
introduces high variance into the policy gradient: the critic has not yet fully integrated the
latest target information before its gradient is used to move the policy.
TD3 therefore updates the actor only once every $K = 2$ critic steps.
This delay ensures the critics are in a more stable and accurate state when the actor gradient
is computed, substantially reducing the variance of policy updates and preventing the actor from
over-adapting to transient critic errors~\cite{fujimoto2018:TD3}.
In addition, the target networks for both actor and critics are updated via the shared Polyak
rule~\eqref{eq:polyak}, ensuring the learning targets change slowly and the training signal
remains consistent across consecutive gradient steps.

TD3 is particularly well-suited for the hockey task.
The environment demands a combination of short-horizon precision---intercepting a fast-moving
puck, executing a well-timed shot---and long-horizon strategic positioning and defense.
This mixture is challenging for actor-critic methods because a single erroneous value estimate in
a critical state can cascade into a series of aggressive, high-variance actions that destabilize
training.
TD3's conservative minimum targets directly bound the magnitude of such errors, target policy
smoothing prevents the actor from locking onto narrow value peaks that may correspond to
unrealistic shooting strategies, and delayed updates ensure the policy only improves once the
critic has formed a reliable local picture of the value landscape.
Together, these mechanisms provide the stable learning dynamic needed to acquire the varied
behavioural repertoire required in the hockey environment.

\subsection{Soft Actor-Critic}
\label{sec:sac}

\ac{SAC} is an off-policy actor-critic algorithm designed for optimizing stochastic policies in
continuous action spaces.
Our implementation follows the framework introduced by~\cite{haarnoja2018soft}, which augments the
standard maximum reward objective with an entropy maximization term~\cite{ziebart2010}.
This approach encourages wider exploration and improves robustness by seeking a policy that
maximizes the trade-off between expected return and entropy:
\begin{equation}
  J(\pi) = \sum_{t=0}^{T}
    \mathbb{E}_{(\mathbf{s}_t,\mathbf{a}_t)\sim\rho_\pi}\!\left[
      r(\mathbf{s}_t,\mathbf{a}_t) + \alpha\,\mathcal{H}\!\left(\pi(\cdot|\mathbf{s}_t)\right)
    \right]
\end{equation}
where $\rho_\pi$ denotes the state-action distribution induced by policy $\pi$ and $\mathcal{H}$
represents the entropy of the policy at state $\mathbf{s}_t$.

The \ac{SAC} algorithm consists of an actor, critic, and an entropy regularization term governed
by a temperature parameter $\alpha$.
The actor is parameterized by a neural network that defines a Gaussian distribution over the
continuous action space by estimating the mean $\mu_\theta(s)$ and standard deviation
$\sigma_\theta(s)$ in its output layer.

To ensure differentiability of the sampling process with respect to the policy parameters, the
reparameterization trick is applied~\cite{haarnoja2018soft, kingma2014auto}.
Instead of sampling actions directly from the Gaussian distribution, actions are expressed as
$a = \mu_\theta(s) + \sigma_\theta(s) \odot \epsilon$ where $\epsilon \sim \mathcal{N}(0, I)$ is
a standard normal random variable.

The sampled actions are passed through a $\tanh$ function and scaled to match the action bounds
of the environment.
To correctly account for this squashing transformation in the entropy term, the log-probability
is adjusted using the determinant of the Jacobian of the $\tanh$ transformation.
The actor is trained on the maximum entropy objective~\cite{haarnoja2018soft}:
\begin{equation}
  J(\pi_\phi) =
    \mathbb{E}_{(\mathbf{s}_t \sim \mathcal{D})}\!\left[
      \mathbb{E}_{(\mathbf{a}_t) \sim \pi_\phi}\!\left[
        \alpha \log \pi_\phi(a_t|s_t) - Q_\theta(s_t, a_t)
      \right]
    \right]
\end{equation}
where for each state sampled from the replay buffer $\mathcal{D}$, an action is sampled from the
actor and the critic estimates the value for each state-action pair.

As presented in the follow-up work~\cite{haarnoja2018algorithms}, we employ two Q-networks for
the critic instead of a single Q-network and a value network as originally proposed.
The soft value function can be implicitly derived from the soft Q-function by combining the
Q-values with the policy entropy term.

During training, two independent Q-networks estimate the value of a state-action pair.
To mitigate overestimation bias, soft policy evaluation is performed using the minimum of the two
Q-values, which provides a more conservative value estimate~\cite{Fujimoto2018Addressing}.
The critic is trained using a mean squared Bellman error loss:
\begin{equation}
  J_Q(\theta_i) = \mathbb{E}_{\mathcal{D}}\!\left[
    \left(Q_{\theta_i}(s_t,a_t) - y_t\right)^2
  \right]
\end{equation}
with target value
\begin{equation}
  y_t = r_t + \gamma\!\left(
    \min_{j=1,2} Q_{\bar{\theta}_j}(s_{t+1},a_{t+1})
    - \alpha \log \pi_\phi(a_{t+1}|s_{t+1})
  \right).
\end{equation}
The target networks $\bar{\theta}_j$ are updated using soft updates with factor $\tau$, which
improves training stability by preventing rapid divergence between online and target critics.

Furthermore, automatic parameter tuning is implemented as introduced in the follow-up
paper~\cite{haarnoja2018algorithms}.
The temperature parameter $\alpha$ is optimized to match a target policy entropy, allowing the
agent to automatically balance exploration and exploitation.
The temperature is updated by minimizing
\begin{equation}
  J(\alpha) = \mathbb{E}_{a_t \sim \pi_\phi}\!\left[
    -\alpha\left(\log \pi_\phi(a_t|s_t) + \bar{\mathcal{H}}\right)
  \right],
\end{equation}
where $\bar{\mathcal{H}}$ denotes the target entropy, typically set to
$-\operatorname{dim}(\mathcal{A})$ as recommended in~\cite{haarnoja2018algorithms}.
The parameter $\alpha$ is optimized using a gradient-based method.


\subsection{Improvements}
\label{sec:improvements}

We implemented a framework based on YAML configuration files to manage and persist training runs,
enabling reproducible experiments and systematic tracking of hyperparameter changes.
All modifications described below are applied on top of both the TD3 and SAC base implementations.


\subsubsection*{Replay Buffer}
Since both implemented algorithms are off-policy, training relies on data stored in a replay
buffer.
We implemented standard \ac{ER}, which improves data efficiency by reusing past interactions and
reducing correlations between subsequent samples.
Transitions are stored as tuples $(s_t, a_t, r_t, s_{t+1})$ in a finite buffer; during training,
batches are sampled uniformly from the buffer.
However, uniform sampling treats all transitions equally, although some experiences may be more
informative for learning.
To address this limitation, we additionally implement \ac{PER}.


\paragraph{Prioritized Experience Replay}
\ac{PER} assigns higher sampling probabilities to transitions with larger temporal-difference (TD) errors, under the assumption that such samples provide more informative learning signals~\cite{Schaul2015PrioritizedER}. Instead of uniform sampling, each transition $i$ is associated with a priority value $p_i = |\delta_i| + \varepsilon$, where $\delta_i$ denotes the TD error and $\varepsilon > 0$ ensures non-zero probability. Transitions are sampled according to 
\begin{equation}   
  P(i) = \frac{p_i^\alpha}{\sum_k p_k^\alpha}, 
\end{equation} where $\alpha$ determines the degree of prioritization ($\alpha = 0$ recovers uniform sampling). Prioritized sampling introduces a bias, corrected via importance-sampling (IS) weights: 
\begin{equation}   
  w_i = \left(\frac{1}{N} \cdot \frac{1}{P(i)}\right)^\beta, 
\end{equation} where $N$ is the buffer size and $\beta$ controls the strength of bias correction~\cite{Schaul2015PrioritizedER}.
To progressively compensate for the bias introduced by prioritized sampling, the importance-sampling exponent $\beta$ is annealed towards $1$ over the course of training, ensuring asymptotically unbiased updates. 
In our implementation, the annealing schedule is controlled via a configurable parameter that specifies the fraction of total training steps over which $\beta$ is linearly increased to its final value of $1$, after which it remains constant. This ensures that updates become fully bias-corrected in the later stages of training.
Efficient sampling and priority updates are implemented using a binary SumTree data structure, enabling both operations in $\mathcal{O}(\log N)$ time.


\subsubsection*{Pink Noise}


\subsubsection*{Reward Shaping}
To enhance learning in the hockey environment, we extend the original reward formulation with additional shaping terms. The environment provides a sparse positive reward for winning as well as a reward related to the agent’s distance to the puck. Furthermore, the environment includes signals for puck contact and puck direction, which are not enabled by default.

In early training stages, we observed that the agent often needed a significant amount of time to start interacting with the puck. To accelerate early learning and encourage active engagement, we introduce a time-penalty term that is applied when the puck remains stationary on the agent’s side of the field. This discourages too passive behavior and promotes faster initiation of play.
To encourage more organized defensive behavior and prevent the agent from leaving its own goal unprotected, we also include a shaping reward based on the distance to the agent’s goal

All shaping components are configurable via the YAML configuration file and can be individually weighted or disabled. This design allows systematic experimentation with different reward combinations.
\subsubsection*{Self-Play}
