\section{Evaluation}\label{sec:eval}
In this section, we evaluate the performance of the implemented algorithms and the proposed improvements across multiple experimental settings. The evaluation is structured in three stages.
First, we assess the general performance and stability of the algorithms in three commonly used benchmark environments.
Second, we analyze training behavior in the hockey environment, with particular focus on the impact of the introduced modifications, including self-play and other algorithmic adjustments. 
Finally, we evaluate the relative playing strength of the resulting agents in a competitive tournament setting. Match outcomes are aggregated and modeled using the Placket-Luce rating system in order to obtain a grounded ranking of agents. This allows a fundamental comparison of different training configurations that goes beyond the performance metrics of a single run.
\subsection{Simple Environments}
We evaluate both \ac{SAC} and \ac{TD3} in three standard benchmark environments: Pendulum, Lunar Lander, and Half Cheetah.  For each algorithm, we compare training with standard \ac{ER} and with \ac{PER} in order to assess the effect of prioritized sampling on learning dynamics and final performance.


\subsection{Hockey Environment}

Own vs. Standard -> trained gegen strong und weak

Plot max 4/ 5 models
TD3 mit PER 
SAC (original vs own)
TD3 (original vs own)


Self Play + Turnier -> Auswahl von model für Turnier

jeweils 1 selfplay (ohne reward shaping)
jeweils 1 gegen strong/ weak trainiert
jeweils reward shaping (mit oder ohne selfplay)
jeweils 1 nur gegen strong trainiert 
basic strong/ weak